Opportunities and precautions for the use of MPAs in bycatch reduction

### 2021-10-12 documents/03_marlin_paper.Rmd
finally back to main

---
title: "Solutions to the potential for unintended conservation consequences of MPAs"
subtitle: Conditions for success and unintended consequences in pelagic MPA design
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::pdf_document2: default
  bookdown::html_document2: default
bibliography: references.bib
params:
  run_name: ["v1.1"]
---

@@

c

\*\* something to point out: increased catches are a paradox, either
come from rebuilding stock or concentrated fishing, but trajectory can
tell you something. not possible for MPAs to increase catches without
decreasing catches first if they are rebuilding? \*\*

Seasonal movemetn is kind of interesting, basically just dilutes the
extremes

less resilient species are both more in need of protection and easier to
accidentally fish harder

@@

@@

in the Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 5.9% of the world's oceans currently in
implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
calling for 30% protection by 2030. The modern science of MPAs was
largely developed around coastal ecosystems, often characterized by
relatively static features such as coral reefs [citation]. However,
large MPAs are increasingly being implemented and considered across
pelagic ecosystems. These pelagic systems are characterized by highly
mobile species and fishing fleets using a diverse mix of static and
dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
also home to combinations of relatively abundant species targeted by
fishing, such as many tuna stocks [@pons2016], and highly threatened
species caught as incidental "bycatch" in targeted fishing operations,
such as oceanic whitetip sharks [@pacoureau2021]. Large pelagic MPAs are
increasingly being proposed specifically to benefit these by-catch
populations while minimizing losses or where possible providing benefits
to fisheries [citation].

In this paper we evaluate the potential conservation and fishery effects
of MPAs within the pelagic context. We focus in particular on the
potential for pelagic MPAs to provide conservation benefits to bycatch
species. While the effect of MPAs on fisheries yields is a continuously
debated topic, the conservation impacts of MPAs are seemingly simpler,
albeit challenging to detect in many cases [@ovando2021a], due in part
perhaps to the perceived obvious nature of their benefits: if you
prevent fishing of a species within an area, there will be more of that
species in that area, and so the conservation benefits of MPAs will
increase with MPA size. \@hastings2017 clearly illustrates this idea.
Using a simple two-species two-patch analytical model, they derive
conditions under which MPAs can solve the "weak-stock" problem in which
a less productive species is caught as btcatch by a fleet targeting a
more valuable and productive stock.

However, as @hastings2017 point out, their results depend on a very
specific set of assumptions, namely that 1) adults are perfectly
sedentary 2) the target and bycatch species overlap perfectly in terms
of habitat and 3) recruits are perfectly mixed throughout the system.
Pelagic species are likely to violate one or more fo these assumptions.
We build on \@hastings2017 to consider the role of MPAs in reducing
bycatch while managing tradeoffs with fisheries under a more resolved
model reflecting the dynamics of pelagic ecosystems and fisheries.

Pelagic ecosystems present both opportunities and challenges for MPAs.
The highly productive and mobile nature of targeted fisheries in these
systems may mean that highly targeted MPAs can be designed that provide
effective conservation at minimal fishery costs. Strategic protection of
key aggregation locations may enhance depleted pelagic fisheries [xx
that Indian ocean study I can't remember right now]. However, the
dynamic nature of pelagic systems also creates a risk of unintended
conservation consequences of MPAs. When an MPA is implemented on former
fishing grounds, fishing fleets have two choices: exit the fishery or
move to new fishing grounds. When MPAs redistribute effort away from one
area and into another, they can accidentally increase fishing effort on
species found within the new fishing grounds [@abbott2012]. This problem
of policy "leakage" is well known in many environmental fields
[@citations] but the general properties of this problem have received
relatively little attention in the theoretical or empirical MPA
literature [see
@abbott2012;@savina2008;[fulton2011a;\@baum2003](mailto:fulton2011a;@baum2003){.email}
for notable exceptions].

This study addresses this gap using a novel bio-economic simulation
model, capable of reflecting several of the key challenges of pelagic
ecosystems, namely high and variable movement rates, mobile fishing
fleets, and heterogeneous species distributions. We first provide a set
of stylized simulations intended to illustrate the potential and
challenges of achieving conservation objectives through pelagic MPAs,
and highlight particularly important drivers and the magnitude and
direction of MPA effects. We then present a more focused case study
based on the Western and Central Pacific Ocean, examining the potential
outcomes of pelagic MPAs on bycatch and target species.

<!-- We fill this gap by examining the general range of conditions… -->

\<!-- \# Assessing the conditions under which this may be a problem
using a generalized simulation environment

General insights about design strategy, effect of varying degrees of
habitat overlap, etc, tradeoffs with fisheries

-   Given that "it depends" nature of the problem, assess the specific
    potential for this problem in a case study

    -   specific insights about design strategy, effect of varying
        degrees of habitat overlap, etc, tradeoffs with fisheries

Our goal here is not to provide tactical recommendations on the design
of blue-water MPAs, but rather to provide a strategic evaluation of the
sensitivity of bycatch populations and targeted fisheries to key
unknowns and design choices in MPA planning. --\>

<!-- The outcome of MPAs for fishing and conservation are highly sensitive to -->

<!-- a wide range of economic and ecological factors. In order to constrain -->

<!-- the possible parameter space, we tune a simulation model based around -->

<!-- best available knowledge of the species and fleets operating in the -->

<!-- areas of the western central pacific ocean (WCPO) while allowing key -->

<!-- unknowns to vary, in particular the nature of adult and larval movement, -->

<!-- the degree of habitat overlap between different species, the economic -->

<!-- incentives facing fishing fleets, and the design strategy used in MPA -->

<!-- placement XX. Most conventional MPA design models are either static, -->

<!-- single species, or both. We present a dynamic, multi-species simulation -->

<!-- model that allows us to cohesively assess the impact of MPAs on a suite -->

<!-- of by-catch and targeted species collectively affected by multiple -->

<!-- fishing fleets.  -->

<!-- We show - something about how when species are homogeneous in space, -->

<!-- conservation gains are basically guaranteed except for constant catch -->

<!-- scenearios, which need to be considered in tuna-type scenarios - When we -->

<!-- introduce heterogeneity in stock structure roughly reflective of -->

<!-- distributions in the WCPO we do / do not begin to see tradeoffs in -->

<!-- conservation outcomes among bycatch species - quantify if there is a -->

<!-- concave function to this - Specific design paradigms are / are not more -->

<!-- / less resilient to this issue - Compare tradeoffs between MPA designs -->

<!-- that minimze internal bycatch tradeoffs (there have to be none if MPA = -->

<!-- 100%) and fishery outcomes -->

<!-- We show that MPAs increased biomass of by-catch species in XX% of -->

<!-- simulations, and decreased them in XX% of simulations. On average a 1% -->

<!-- increase in by-catch species resulted in a XX% change in fishery -->

<!-- catches... Outcomes were most sensitive to XX and XX, and least -->

<!-- sensitive to XX and XX. We do not suggest that these same results would -->

<!-- hold for other systems, but rather demonstrate the importance / -->

<!-- resilience of MPAs to key uncertain variables. In particular for pelagic -->

<!-- ecosystems, our work demonstrates the importance of considering the -->

<!-- collective impacts of space, vulnerability, life history, and fleet -->

<!-- dynamics in MPA design for conservation. -->

@@

@@

### Fleet Dynamics

The key feature of `marlin` for the purposes of this project is the
ability to capture technical interactions across species. While marlin
does not simulate species interactions, fishing fleets in marlin are
capable of targeting and affecting multiple species simultaneously. This
allows us to capture for example the potential impacts of displacement
of effort from one area into another across multiple species.

A fishing fleet or *metier* is defined in `marlin` by a total pool of
effort that can have different catchability, prices, and selectivity for
individual species within a simulation. For example, both a longline and
purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack
(*Katswamus pelamus*), but the longline fleet is much more likely to
catch bigeye than skipjack, and *vice versa*. The fishing fleet then
generates catches, revenues, and profits from individual species that it
targets, and makes decisions based on its total objective function
across all species.

@@

$$
 R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,f}q_{s,f}E_{t,p,f}}{F_{t,p,s,a}} \times Y_{t,p,s,a}
$$

where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
at age, *q* is catchability , *F* is total fishing mortality, and *Y*
are total catches.

@@

$$
F_{t,p,s,a} = \sum_f^F\alpha_{a,f}q_{s,f}E_{t,p,f}
$$

and

@@

This means that each species experiences a total mortality at age in a
time step in a patch, individual fractions of which are portioned off as
catches and subsequently revenues for each fleet. Given revenues, we
then calculate profits \$\\phi\$ per

$$
\phi_{t,p,f} = R_{t,p,f} - c_f(E_{t,p,f} + 1)_{t,p,f}
^{1.3}$$

where *c* is cost per unit effort and, the 1.3 exponent is in the manner
of @costello2016a . The addition of one to the total effort is to
constrain the value to be greater than 1, preventing undesired behavior
where fishing effort actually gets cheaper if total effort is less than
1.

The important thing to note here then is that the revenues and profits
for a given fleet are a function of the prices and catchability of
individual species, but that mortality itself is independent of price.
So, fleet that has selectivity and catchability greater than zero can
exert fishing mortality on a species even if it obtains no value from
that species.

The fleet then decides how to allocate its effort in space based on one
of four possible spatial allocation strategies. The ideal free
distribution (IFD) (citation) is the standard method for distributing
fishing fleets in spaces. However, the IFD presents a number of
complications for our study. First, in order to simplify the assessment
of the impacts of different fishing mortality rates we simulate
scenarios with a fixed total pool of effort. As MPAs change the spatial
distribution of profits though, the total amount of effort required to
satisfy the IFD may not equal the total amount of effort required by the
model. Second, even if we were to implement the IFD for a single fleet,
the model is capable of tracking multiple fleets at once. This means
that the IFD for a single fleet would be a game-theoretic outcome of the
actions of each of the individual fleet conditional on the actions of
all other fleets. While possible to implement this would slow down our
model runs to the point of practically preventing large-scale evaluation
of the effects of MPAs on bycatch.

As such we explored a series of "next best" fleet distribution
algorithms. While not the IFD in any individual time step, over time
they start to approximate the IFD (check this) as each assumes that
fleets base their decisions for the current time step on the outcomes in
the prior timestep, meaning that the impacts of the actions of other
fishing fleets are somewhat internalized. The four possible fleet
distribution algorithms are

@@

### Population Model

Lots of equations. Pretty standard age-structured population models,
extra bells and whistles that need explaining

-   Design of movement and habitat

    -   Movement is a product of distance and habitat, such that
        critters will all else equal move to the closest patches with
        the best habitat

-   Ontogenetic shifts

-   Nature of fleet tuning (shared effort different q per species)

Fleets exert a given amount of effort in a patch each time step. The
catch resulting from that effort is given by the Baranov [@baranov1918]
equation

$$C_{f,s,p,t} = \frac{q_{f,s}E_{f,p,t}}{q_{f,s}E_{f,p,t} + m_s} \times \sum_{a=0}^{A}B_{s,p,t,a} \times (1 - e^{-(q_{f,s}E_{f,p,t}S_a + m_s)}) $$

Where $C$ is catch by fleet $f$ of species $s$ in patch $p$ in timestep
*t*. $E$ is total effort by that fleet in a patch and time step, $m$ is
natrual mortality for a given species, $B$ is biomass at age *a*, and
$S$ is selectivity.

$q$ is the catchability coefficient for fleet $f$ and species $s$.

Fleets distribute themselves in space according to revenue per unit
effort

$$R_{f,p,t} = \frac{\sum_{s=1}^{S}v_{f,s}C_{f,s,p,t}}{E_{f,p,t}}$$ Where
$v$ is the price per kg of each species. Bycatch in the model is defined
through the price. An unregulated species with no commercial value would
have $v = 0$, whereas a regulated and prohibited species might actually
have a negative price.

Effort per fleet is set a a constant of 1 per patch, and the model then
tunes $q_{f,s}$ such that given all the other parameters of the model
the fishing mortality rate for each species matches the best available
level for that species. This is what allows the model to capture bycatch
dynamics: A fleet can distribute itself purely based on revenues from
target species, but through the $q$ parameter still exert fishing
mortality on other species, even if that species has a price of zero, if
those bycatch species inhabit the same patches as the more valuable
target species.

###

### MPA Design Strategies

MPA design is governed by four general parameters: the number of cells
to place inside an MPAs, and the design strategy used to place the MPAs,
the species considered in the MPA planning process, and the accuracy of
the information used in the planning process.

@@

For the "depletion" strategy, MPA locations are prioritized based in
proportion to the spawning stock biomass (SSB) of the most depleted
species. So, patches with high amounts of highly depleted species are
prioritized over patches with high biomass of lightly depleted species.

For the "rate" strategy, MPAs are placed in based on the
depletion-weighted catch relative to the total catch in a patch. So,
patches where with high rates of catch of depleted species relative to
totcal catch are prioritized.

For the "avoid_fishing" strategy, MPAs locations are prioritized
inversely to fishery catches. So, patches with low catches are
prioritized over patches with high catches.

For the "target_fishing" strategy, MPAs locations are prioritized
proportionally to fishery catches. So, patches with high total catches
are prioritized over patches with high catches.

For the "area" strategy, we simply assign patches in sequential order
until the desired total MPA size is reached.

The information used to inform each MPA design strategy (except area)
can include a multiplicative observation error term $\sigma_{obs}$ where

$$information \sim lnorm(data,\sigma_{obs}) $$

For example, for the depletion scenario, $\sigma_{obs}$ controls the
accuracy of the spatial depletion data used to prioritize MPA placement.

@@

### Species & Fleets

Each simulation considers `n_distinct(marlin_inputs$scientific_name)`
species comprised of the top 12 species by landings reported in the tuna
fisheries of the Western and Central Pacific Ocean (WCPO). Rough Life
history parameters are pulled from `FishLife` [@thorson2020]. For
teleost fishes, we assumed that fecundity was proportional to
weight-at-age times the maturity ogive. For cartilaginous species, we
adopt a fixed fecundity relationship of XX pups per mature individual XX
check on what I actually implemented here XX. The unfished spawning
stock biomass for each of the included species was tuned in the model to
reflect best available estimates from stock assessments.

Each simulation includes a longline and a purse-seine fleet. Each
species can have a separate price (USD/kg) and selectivity ogive for
each fleet, as dictated by the literature.

### Species Overlap

@@

### Seasonal Migration

For this paper `marlin` is set to run with two seasons per year. When
seasonal movement is turned on, animals use one habitat matrix in season
one, and the inverse of that habitat matrix in season 2.

## Case Study Application

>>>>

---
title: "Opportunities and precautions for the use of MPAs in bycatch reduction"
author:
    - Daniel Ovando
    - Darcy Bradley
    - Lennon Thomas
    - Echelle Burns
date: "`r Sys.Date()`"
output: 
  bookdown::word_document2:
    reference_docx: template.docx
  bookdown::html_document2: default
  bookdown::pdf_document2: default
header-includes:
  - \usepackage{setspace}\doublespacing
  - \usepackage{lineno}\linenumbers
bibliography: references.bib
params:
  run_name: ["v1.1"]
---

@@

<!-- \*\* something to point out: increased catches are a paradox, either -->
<!-- come from rebuilding stock or concentrated fishing, but trajectory can -->
<!-- tell you something. not possible for MPAs to increase catches without -->
<!-- decreasing catches first if they are rebuilding? \*\* -->

<!-- Seasonal movement is kind of interesting, basically just dilutes the -->
<!-- extremes -->

<!-- less resilient species are both more in need of protection and easier to -->
<!-- accidentally fish harder -->

@@

<!-- MAKE IT MORE ABOUT BYCATCH-->

@@

Marine protected areas (MPAs) are a growing part of the marine
conservation toolbox, with 5.9% of the world's oceans currently in
implemented MPAs [@grorud-colvert2021], and bodies such as the IUCN
calling for 30% protection by 2030. One of the potential benefits of
MPAs is their ability to provide protection to broad segments of an
ecosystem at once, rather than requiring targeted management based on
individual species that may be unfeasible in many parts of the world
lacking in data or fisheries management capacity. Their ability to
provide protection across a range of species with one policy has led them to be suggested as a tool to reduce problems around weak stock or "bycatch"
species [@hastings2017]. We use the term "bycatch" here to generally refer to the
capture of species within a fishery primarily targeting a different set
of species.

While the effect of MPAs on fisheries yields is a continuously debated
topic, the conservation impacts of MPAs are seemingly simpler, albeit
challenging to detect in many cases [@ovando2021a], due in part perhaps
to the perceived obvious nature of their benefits: if you prevent
fishing of a species within an area, there will be more of that species
in that area, and so the conservation benefits of MPAs will increase
with MPA size. @hastings2017 clearly illustrates this idea. Using a two-species two-patch analytical model, they derive conditions
under which MPAs can solve the "weak-stock" problem in which a less
productive species is caught as bycatch by a fleet targeting a more
valuable and productive stock.

However, as @hastings2017 point out, their results depend on a very
specific set of assumptions, namely that 1) adults are perfectly
sedentary 2) the target and bycatch species overlap perfectly in terms
of habitat and 3) recruits are perfectly mixed throughout the system. In
this paper We build on @hastings2017 to consider the role of MPAs in
reducing bycatch while managing tradeoffs with fisheries under a more
resolved model.

Specifically, the potential benefits of MPAs as a bycatch management are
complicated by two key issues: effort displacement and technical
interactions. One of the primary questions around MPAs is to what extent
do MPAs remove versus displace fishing effort. If MPAs remove fishing
effort, then their direct effects on target and bycatch species can only
be positive. If however MPAs displace fishing effort evaluating their
potential impacts becomes more complicated. The net conservation impact
of MPAs becomes a balancing act between the effects of reduced fishing
inside protected borders and increased fishing intensity outside MPAs.
The outcomes of this displacement can be further complicated by
technical interactions, i.e. the degree to which fishing effort for one
species affects another, particularly when species have markedly
different habitat distributions. For example, @abbott2012 demonstrated
that an MPA designed to protect red king crab (*Paralithodes
camtschaticus*) caused an unintended increase in bycatch of Pacific
halibut (*Hippoglossus stenolepis*) by the displaced bottom-trawl
fishery. Policy leakage is not a unique feature to MPAs, e.g. XX, but
the potential impacts of MPAs on target and bycatch species taking into
account the potential for these leakage effects has received relativelly
little generalized attention in the literature [see @abbott2012;
@savina2008; @fulton2011a; @baum2003 for notable exceptions].

This study addresses this gap using a novel bio-economic simulation
model that allows for efficient evaluation of the effects of MPAs on
target and bycatch species across a range of ecological and economic
dynamics, critically accounting for the potential impacts of effort
displacement and technical interactions. We use this model first to
explore general conditions that lead to positive and negative outcomes
for target and bycatch species from MPAs. We then present a case study
tuned to reflect the dynamics of the tuna, billfish, and shark
populations and fisheries of the Western and Central Pacific Ocean
(WCPO) to explore the value of optimal MPA design in managing tradeoffs
between maximizing bycatch conservation and minimizing economic impacts.
We chose the WCPO as a case study as the modern science of MPAs was
largely developed around coastal ecosystems, often characterized by
relatively static features such as coral reefs [citation]. However,
large MPAs are increasingly being implemented and considered across
pelagic ecosystems. These pelagic systems are characterized by highly
mobile species and fishing fleets using a diverse mix of static and
dynamic ocean features [@game2009; @lubchenco2003]. These ecosystems are
also home to combinations of relatively abundant species targeted by
fishing, such as many tuna stocks [@pons2016], and highly threatened
species caught as incidental "bycatch" in targeted fishing operations,
such as oceanic whitetip sharks [@pacoureau2021]. These dynamics,
combined with the increased policy interest in large MPAs covering
expanses of blue-water ecosystems, make pelagic systems such as the WCPO
an ideal case study in bycatch management through MPAs. The goal of this
paper is to provide a strategic evaluation of the sensitivity of bycatch
populations and targeted fisheries to key unknowns and design choices in
MPA planning.

@@

Below we provide a summary of the `marlin` model, as well as details of
the generalized simulation exercises and case study analyses.

@@

#### Fleet Dynamics

The key feature of `marlin` for the purposes of this project is the
ability to capture technical interactions across species. While marlin
does not simulate species interactions, fishing fleets in `marlin` are
capable of targeting and affecting multiple species simultaneously. This
allows us to capture for example the potential impacts of displacement
of effort from one area into another across multiple species.

A fishing fleet or *metier* is defined in `marlin` by a total pool of
effort that can have different catchability, prices, and selectivity for
individual species within a simulation. For example, both a longline and
purse seine fleet may capture bigeye (*Thunnus obesus*) and skipjack
(*Katswamus pelamus*), but the longline fleet is much more likely to
catch bigeye than skipjack, and *vice versa*. Each fishing fleet
generates catches, revenues, and profits from individual species that it
targets, and makes decisions based on its total objective function
across all species.

@@

$$
 R_{t,p,f} = \sum_s^S \sum_a^A \Pi_{s,f}\frac{\alpha_{a,s,f}q_{s,f}E_{t,p,f}}{F_{t,p,s,a}} \times Y_{t,p,s,a}
$$

where *p* is patch, *s* is species, *a* is age, $\alpha$ is selectivity
at age, *q* is catchability , *F* is total fishing mortality, $\Pi$ is
price per unit weight, and *Y* is total catches.

@@

$$
F_{t,p,s,a} = \sum_f^F\alpha_{a,s,f}q_{s,f}E_{t,p,f}
$$

Selectivity $\alpha$ is modeled through a logistic form per

```{=tex}
\begin{equation}
\alpha_{a,s,f}=\frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - l^{sel}_{s,f}}{\delta^{sel}_{s,f}}})}
  (\#eq:sel)
\end{equation}
```
where $l$ is length at age, $l^{sel}$ is the length at 50% selectivity,
and $\delta$ is the difference between the length at 50% selectivity and
the length at 95% selectivity.

*z* is total mortality given by

@@

This means that each species experiences a total mortality at age in a
time step in a patch, individual fractions of which are portioned off as
catches and subsequently revenues for each fleet. Given revenues, we
then calculate profits $\phi$ per

$$
\phi_{t,p,f} = R_{t,p,f} - c_f(E_{t,p,f} + 1)_{t,p,f}
^{1.3}
$$

where *c* is cost per unit effort and, the 1.3 exponent is in the manner
of @costello2016a . The addition of one to the total effort is to
constrain the value to be greater than 1, preventing undesired behavior
where fishing effort actually gets less expensive if total effort is
less than 1.

The important feature here then is that the revenues and profits for a
given fleet are a function of the prices and catchability of individual
species, but that mortality itself is independent of price. So, a fleet
that has selectivity and catchability greater than zero can exert
fishing mortality on a species even if it obtains no value from that
species.

Each fishing fleet then decides how to allocate its effort in space
based on one of four possible spatial allocation strategies. The ideal
free distribution (IFD) is the standard method for distributing fishing
fleets in spaces (see @gillis2003 and references therein). However, the
IFD presents a number of complications for our study. First, in order to
simplify the assessment of the impacts of different fishing mortality
rates we simulate scenarios with a fixed total pool of effort. As MPAs
change the spatial distribution of profits though, the total amount of
effort required to satisfy the IFD may not equal the total amount of
effort required by the model. Second, even if we were to implement the
IFD for a single fleet, the model is capable of tracking multiple fleets
at once. This means that the IFD for a single fleet would be a
game-theoretic outcome of the actions of each of the individual fleet
conditional on the actions of all other fleets. While possible to
implement this would slow down our model runs to the point of
practically preventing large-scale evaluation of the effects of MPAs on
bycatch.

As such we explored a series of "next best" fleet distribution
algorithms. While not the IFD in any individual time step, over time
they start to approximate the IFD, as each assumes that fleets base
their decisions for the current time step on the outcomes in the prior
timestep, meaning that the impacts of the actions of other fishing
fleets are somewhat internalized. The four possible fleet distribution
algorithms are

@@

Revenue based spatial distribution is not likely to be very realistic .
However, due to the complexity of parameterizing cost functions fleet
dynamics are often evaluated based on yield or revenue alone, and so we
include those scenarios here to evaluate the potential implications of
this choice. The decision on whether to allocate the fleet based on
absolute or relative (per unit effort) is more complex. When effort
represents the actions of separate and individual fishing actors (e.g.
independent fishing vessels), a per-unit-effort strategy may be more
realistic, in which fishers distribute themselves based on the expected
catch of their individual efforts [@hilborn1987]. Conversely, a system
defined by a sole owner seeking to maximize total profits might be
better represented by a fleet model based on total profits

#### Population Model

The underlying population model used is an age structured single-species
model in the manner of @ovando2021a.

For the population model, numbers *n* at time *t* for age *a* are given
by

```{=tex}
\begin{equation}
n_{t,p,s,a}=\begin{cases}
      = BH(ssb_{t-1,p,s,a}) & \text{if $a = 1$}\\
     = n_{t-1,p,s,a-1}e^{-(z_{t-1,p,s,a-1}}, & \text{if $1< a < max(age)$}\\
     =  n_{t-1,p,s,a}e^{-(z_{t-1,p,s,a})} + n_{t-1,a-1}e^{-(z_{t-1,p,s,a-1})}, & \text{if $a = max(a)$}
  \end{cases}
  (\#eq:pop)
\end{equation}
```
where *BH* is the Beverton-Holt recruitment function, *ssb* is
spawning-stock-biomass.

where $l_a$ is the mean length at at age, $l_sel$ is the length at which
on average 50% of individuals are selected by the fishery, and
$\delta_{sel}$ are the additional units of length at which on average
95% of fish are selected by the fishery.

*ssb* is calculated by converting age to mean length, calculating weight
at age, maturity at age, and then calculating spawning stock biomass as
the sum of spawning potential at age in a given time step.

```{=tex}
\begin{equation}
  l_{a,s} = l\infty_s\left(1 - e^{-k_s(a - a0_{s})}\right)
  (\#eq:len)
\end{equation}
```
Biomass *b* at age is then given by

```{=tex}
\begin{equation}
  b_{a,s} = wa_s \times l_{a,s}^{wb_s}
  (\#eq:weight)
\end{equation}
```
and maturity *mat* is calculated as

```{=tex}
\begin{equation}
  mat_{a,s} = \frac{1}{(1 + e^{-log(19)\times\frac{l_{a,s} - lmat_s}{{\delta}mat_s}}}
  (\#eq:mat)
\end{equation}
```
where $l_{mat}$ is the length at which on average 50% of individuals are
sexual maturity, and $\delta_{mat}$ is the units of length beyond
$l_{mat}$ at which on average 95% of fish are sexually mature.

Spawning stock biomass at time *t* is then calculated as

```{=tex}
\begin{equation}
  ssb_{t,p,s} = \sum_{a=1}^Aw_{a,s,t}^{\gamma_s} mat_{t,a} n_{t,p,s,a}
    (\#eq:ssb)
\end{equation}
```
##### Recruitment {.unnumbered}

Recruitment follows Beverton-Holt dynamics. We do however allow for
three variants in the timing of density dependence:

1.  Global density dependence: Density dependence is a function of the
    sum of spawning biomass across all patches, and recruits are then
    distributed according to habitat quality

```{=tex}
\begin{equation}
n_{t,p,s,a = 1} = \left(\frac{0.8{\times}\sum_{p=1}^P{r0_{p}}\times{h}\times\sum_{p=1}^P{ssb_{t-1,p}}}{0.2\times{\sum_{p=1}^Pssb0_p}\times(1 - h)+(h - 0.2)\times{\sum_{p=1}^Pssb_{t-1,p}}}\right)\times {hab_p}
  (\#eq:dd2)
\end{equation}
```
where *hab* is a vector of recruitment habitat quality by patch degined
by

$$hab_p = r0_p / \sum_p^P{r0_p}$$ 2. Local density dependence: Density
dependence occurs independently in each patch and recruits are retained
in their home patch.

```{=tex}
\begin{equation}
  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)
  (\#eq:dd1)
\end{equation}
```
3.  Local density dependence then disperse: Density dependence occurs
    independently in each patch and recruits are then dispersed.

```{=tex}
\begin{equation}
  n_{t,p,s,a = 1,p} = \left(\frac{0.8{\times}r0_{s,p}\times{h_s}\times{ssb_{t-1,p,s}}}{0.2\times{ssb0_{p,s}}\times(1 - h_s)+(h_s - 0.2)\times{ssb_{t-1,p}}}\right)\times \boldsymbol{d^l} \times \epsilon_t
  (\#eq:dd3)
\end{equation}
```
where **d^l^** is the recruitment movement matrix

4.  Post-dispersal density dependence: Larvae are distributed throughout
    the system, and then density dependence occurs based on the density
    of adult biomass at the destination patch.

```{=tex}
\begin{equation}
  larv_{t,p} = ssb_{t-1} \times\boldsymbol{d^l}
  (\#eq:larvmove)
\end{equation}
```
```{=tex}
\begin{equation}
n_{t,a = 1,p} = \left(\frac{0.8{\times}r0_{p}\times{h}\times{larv_{t,p}}}{0.2\times{ssb0_p}\times(1 - h)+(h - 0.2)\times{larv_{t,p}}}\right) \times \epsilon_t
  (\#eq:dd4)
\end{equation}
```
##### Dispersal {.unnumbered}

Dispersal in the model is broken into two components: adults and
recruits Both assume a Gaussian dispersal kernel of the form

```{=tex}
\begin{equation}
m_{s,p_i,p_j}=\frac{1}{\sqrt{2\pi\sigma_{s,p_i}^2}}e^{-\frac{d_{p_i,p_j}^2}{2\sigma_{s,p_i}^2}}
  (\#eq:move)
\end{equation}
```
where *i* is the source patch, *j* is the destination patch, *d* is the
distance between patches *i* and *j* , and $\sigma_s*$is the movement
rate, in units of patches.

Along with distance, net movement is defined by habitat quality. Habitat
quality defines the quailty for adults or larvae of a given patch. The
net probability of movement from patch *i* to patch *j* is defined then
by the product of distance and habitat quality, normalized to sum to one
from each patch, such that all else being equal organisms will move the
most to the closest patch with the best habitat. We allow for seasonal
movement in the model, where years are broken into two seasons. Movement
in season two is defined as the opposite of movement in season one,
intended to reflect the behavior of seasonal migrations.

## Rule-of-thumb MPA Design Strategies

MPA design is governed by two general parameters: the number of cells to
place inside an MPAs, and the design strategy used to place the MPAs.

@@

1.  Depletion: MPA locations are prioritized in proportion to the
    spawning stock biomass of the most depleted species pre-MPA. So,
    patches with high spawning biomass of highly depleted species are
    prioritized over patches with high spawning biomass of lightly
    depleted species.

2.  Rate: MPAs are placed in based on the pre-MPA depletion-weighted
    catch relative to the total catch in a patch. So, patches with high
    rates of catch of depleted species relative to total catch are
    prioritized.

3.  Avoid_fishing: MPAs locations are prioritized inversely to fishery
    catches. So, patches with low catches are prioritized over patches
    with high catches.

4.  Target_fishing: MPAs locations are prioritized proportionally to
    fishery catches. So, patches with high total catches are prioritized
    over patches with low catches.

5.  Area: For the "area" strategy, we simply assign patches in
    sequential order until the desired total MPA size is reached.

<!-- The information used to inform each MPA design strategy (except area) -->

<!-- can include a multiplicative observation error term $\sigma_{obs}$ where -->

<!-- $$information \sim lnorm(data,\sigma_{obs}) $$ -->

<!-- For example, for the depletion scenario, $\sigma_{obs}$ controls the -->

<!-- accuracy of the spatial depletion data used to prioritize MPA placement. -->

## MPA Optimization

Our MPA optimization algorithm seems to create an MPA network that
maximizes an objective function, given by

$$obj_p = \Phi \times scale(\sum_f^F \boldsymbol{\phi_{f,p}}) + (1 - \Phi)  scale(\boldsymbol{\sum_s^S\frac{ssb_s}{ssb0_{s,p}}})$$

Where *scale* rescales the values to be between 0 and 1 such that the
weight parameter is comparable between vectors. The algorithm works by
first assigning every patch in the system to an MPA. For the first
round, the algorithm then calculates the equilibrium objective function
value resulting from sequentially removing each individual patch from
the MPA network. The cell whose removal from the MPA produces the
highest objective function value is then removed from the MPA network,
and the marginal value of the objective function each of the remaining
cells from the last steps are then stored. For the next steps, we
sampled 25% of the remaining MPA cells with probability proportional to
their marginal contribution to the objective function in the prior
iteration, selecting the cell whose removal produces the highest total
objective function as the cell to remove from the network, and updateing
the marginal value of the sampled-but-not-selected cells. This in
essence is a sampling-importance-resampling algorithm. The algorithm
continues until no more cells are left in the MPA. This then provides
for a given objective weighting value $\Phi$ the near-optimal MPA
network at every MPA size from 0 to 100%. We say "near" as they is some
path dependency here, in that for example the network that maximizes the
objective function with 40% of the MPAs does not necessarily contain the
optimal network at 30%. However, the reversed nature of the algorithm,
starting at 100% MPA coverage and working better, produces a more stable
network with regards to this problem than starting at 0% and growing to
100%.

@@

Each iteration uses one of three species: A stylized tuna modeled after
skipjack tuna (), a stylized marlin modeled after (), and a stylized
shark modeled after (). Each species has all life history parameters
pulled from FishLife (@thorson2020), with the exception of steepness, set
to XX, and fecundity. For the tuna and marlin species, fecundity is
assumed to scale with weight modified by the hyperallometric exponent.
For the shark species, we assume a constant value of six pups per year
at all ages greater than the age at maturity. We assume that tunas have
an unfished biomass of XX, marlins XX, and sharks XX, to reflect a
large, medium, and small population.

For the generalized experiments we only include one fishing fleet. We
asign a price per KG of \$XX for tuna, intended to reflect a low value
high volume fishery, \$XX for marlin for a higher value low volume
fishery, and \$0 for sharks to reflect at true bycatch species. We
assign a selectivity of 60% of length at maturity for skipjack to allow
for some overfishing when F/M is high enough, 70% for mariln, and 90%
for shark.

We ran `r nrow(state_experiments)` state experiments, varing the state
of nature experienced by a simulation. The candidate states include

| State                 | Distribution          | Description                                     |
|-----------------------|-----------------------|-------------------------------------------------|
| $\sigma_{centroid}$   |                       | Standard deviation of habitat centroid          |
| $\sigma_{habitat}$    |                       | Standard deviation around centroid              |
| Spatial catchability  |                       | Catchability varies in sapce                    |
| Spatial allocation    |                       | Fleet allocation strategy                       |
| F/M                   |                       | Fishing mortality relative to natural mortality |
| Adult movement rate   |                       |                                                 |
| Recruit movement rate |                       |                                                 |
| Hyperallometry        |                       | Presence of hyperallometric fec                 |
| DD timing             |                       | Timing of density dependence                    |
| Seasonal Movement     | $\subset[TRUE,FALSE]$ | Seasonal Movement Present                       |

For each of these `r nrow(state_experiments)` state experiments, we then
apply `r nrow(placement_experiments)` placement experiments defined by
an MPA size, an MPA placement strategy, and ...XX

@@

## WCPO Case Study Application

The purpose of the case study is to assess the value of optimization.
Given the time constraints of optimization, rather than iterating over
the `r nrow(state_experiments)`, we fix the values from that table to
values that best reflect the state of
`r n_distinct(experiment_obj$critter)` species reported as caught in the
WCPO.

Sigh. lots of details to add in here. This is the worst.
